\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}

### 2022-03-11 k-planar.tex
comments about framed graphs

\begin{thm}
\label{1-planar}
    Every $d$-framed graph is contained in $H\boxtimes P\boxtimes K_{3\lfloor d/2\rfloor + d -3}$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
\end{thm}

@@

\appendix
\section{New Stuff}

@@

\begin{thm}
\label{CliqueFaceProductStructure}
Let $G$ be a 2-connected plane multigraph, in which every facial cycle has length in $\{3,\dots,d\}$ for some fixed integer $d\geq 3$. Then $\widehat{G}$ is contained in $H \boxtimes P \boxtimes K_{ d  \floor{d/2} }$ for some graph $H$ with $\tw(H) \leq 3$ and for some path $P$.
\end{thm}

@@

\begin{proof}[Proof of \cref{CliqueFaceProductStructure}]
Let $G$ be a 2-connected plane multigraph, in which every facial cycle has length in $\{3,\dots,d\}$. Let $C$ be the outer-cycle of $G$. Let $G^+$ be obtained from $G$ by adding a new vertex $r$ adjacent to every vertex on the outerface of $G$. So $G^+$ is a 2-connected plane multigraph and $G=G^+\blah{C}$. Let $T$ be a BFS spanning tree of $G^+$ rooted at $r$. Let $\mathcal{L}$ be the corresponding BFS layering. By \cref{NewLemma} \note{DW}{with improved (ii)}, $G$ has a partition $\PP$ into tripods with respect to $T$ such that $G / \PP$ has a tree-decomposition $\mathcal{T}$ of width at most 3 such that for every facial cycle $F$ of $G$ there is a bag $B$ of $\mathcal{T}$ such that every part of $\PP$ that intersects $F$ is in $B$. Since $G$ is a spanning subgraph of $\widehat{G}$, we may consider $\PP$ to be a partition of $\widehat{G}$, and $G/\PP \subseteq \widehat{G}/\PP$. Consider an edge $XY$ of $\widehat{G}/\PP$. By the definition of $\widehat{G}$, there is a facial cycle $F$ of $G$ such that both $X$ and $Y$ intersect $F$. By assumption, there is a bag $B$ of $\mathcal{T}$ such that every part of $\PP$ that intersects $F$ is in $B$. In particular, $X,Y\in B$. Hence $\TT$ is also a tree-decomposition of $\widehat{G}$. Each tripod in $\PP$ has at most $d$ vertices in each layer in $\mathcal{L}$. For each edge $vw$ of $\widehat{G}$, if $v\in L_i$ and $w\in L_j$ then $|i-j|\leq\floor{\frac{d}{2}}$. Grouping blocks of $\floor{\frac{d}{2}}$ consecutive layers gives a layering of $\widehat{H}$ with layered width $\floor{\frac{d}{2}}d$ with respect to $\PP$. By \cref{PartitionProduct}, $\widehat{G}$ is contained in $H \boxtimes P \boxtimes K_{ d  \floor{d/2} }$ where $H:=\widehat{G}/\PP$, which has treewidth 3.
\end{proof}

>>>>

\begin{thm}
\label{framed}
    Every $d$-framed graph is contained in $H\boxtimes P\boxtimes K_{3\lfloor d/2\rfloor + d -3}$ for some planar graph $H$ with treewidth at most 3 and for some path $P$.
\end{thm}

\note{DW}{I think \cref{framed} should be proved explicitly and \cref{1-planar} should be a corollary, especially because \cref{NewMapGraph} is another corollary. This section should be about framed graphs, outside of \cref{sec-k-planar}. I suggest we make it a subsection of 
\cref{Structure} where we introduce framed graphs, say they can be described by $(2,\binom{d}{2})$-shortcut systems, so the main theorem is applicable, which gives a treewidth bound of $\binom{5}{3}-1=9$, but using  \cref{framed} we get a treewidth bound of 3, which is optional. Then in latter sections we give two applications, for 1-planar graphs and for $d$-map graphs. }

@@

\appendix
\section{Framed Graphs}

\note{DW}{Use the language of framed graphs}

@@

\begin{thm}
\label{CliqueFaceProductStructure}
Let $G$ be a 2-connected plane multigraph, in which every facial cycle has length in $\{3,\dots,d\}$ for some fixed integer $d\geq 3$. Then $\widehat{G}$ is contained in $H \boxtimes P \boxtimes K_{ d +3( \floor{d/2}-1) }$ for some graph $H$ with $\tw(H) \leq 3$ and for some path $P$.
\end{thm}

@@

\begin{proof}[Proof of \cref{CliqueFaceProductStructure}]
Let $G$ be a 2-connected plane multigraph, in which every facial cycle has length in $\{3,\dots,d\}$. Let $C$ be the outer-cycle of $G$. Let $G^+$ be obtained from $G$ by adding a new vertex $r$ adjacent to every vertex on the outerface of $G$. So $G^+$ is a 2-connected plane multigraph and $G=G^+\blah{C}$. Let $T$ be a BFS spanning tree of $G^+$ rooted at $r$. Let $\mathcal{L}=(L_0,L_1,\dots)$ be the corresponding BFS layering. By \cref{NewLemma}, $G$ has a partition $\PP$ into tripods with respect to $T$ such that $G / \PP$ has a tree-decomposition $\mathcal{T}$ of width at most 3 such that for every facial cycle $F$ of $G$ there is a bag $B$ of $\mathcal{T}$ for which every part of $\PP$ that intersects $F$ is in $B$. Since $G$ is a spanning subgraph of $\widehat{G}$, we may consider $\PP$ to be a partition of $\widehat{G}$, and $G/\PP \subseteq \widehat{G}/\PP$. Consider an edge $XY$ of $\widehat{G}/\PP$. By the definition of $\widehat{G}$, there is a facial cycle $F$ of $G$ such that both $X$ and $Y$ intersect $F$. By assumption, there is a bag $B$ of $\mathcal{T}$ such that every part of $\PP$ that intersects $F$ is in $B$. In particular, $X,Y\in B$. Hence $\TT$ is also a tree-decomposition of $\widehat{G}/\PP$. For each edge $vw$ of $\widehat{G}$, if $v\in L_i$ and $w\in L_j$ then $|i-j|\leq\floor{\frac{d}{2}}$. Grouping blocks of $\floor{\frac{d}{2}}$ consecutive layers in $\mathcal{L}$ gives a layering $\mathcal{L}'$ of $\widehat{G}$. Observe that each tripod in $\PP$ has at most $d+3( \floor{\frac{d}{2}}-1)$ vertices in each set of $\floor{\frac{d}{2}}$ layers of $\mathcal{L}$. Hence $\mathcal{L}'$ has layered width $d+3( \floor{\frac{d}{2}}-1)$ with respect to $\PP$. By \cref{PartitionProduct}, $\widehat{G}$ is contained in $H \boxtimes P \boxtimes K_{ d + 3( \floor{d/2} -1 )}$ where $H:=\widehat{G}/\PP$, which has treewidth 3.
\end{proof}
